---
title: "Estimating Marine Mammal Abundance"
subtitle: "FISH 497"
author: "Zoe Rand"
format:
  revealjs:
    embed-resources: true
    theme: solarized
editor: visual
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(sf)
library(Distance)
bw_dat<-read_csv("data/simulated_bw.csv")
```

## Getting started

-   Install and load `tidyverse`, `sf` and `Distance` packages

-   Unzip the `data` folder and place in your root directory

## Instructions {.smaller}

We will be using data from a simulated line-transect survey for blue whales off the coast of California, Oregon, and Washington.

1\) Read in your data (`simulated_bw.csv`) and look at it. What kind of information is in each of the columns?

2\) Plot the locations of each detection on top of the tracklines. You will need to read in the `tracklines.shp` and `region.shp` files using `read_sf()`. You can plot these using `geom_sf()`.

3\) Make a histogram of your detection distances.

4\) Run three distance sampling models using the `ds()` function in the `Distance` package. You will need to select a detection function for each model: half-normal = `"hn"`, hazard rate = `"hr"`, and uniform = `"unif"`. You will also need to set `convert_units = convert_units("meter", "meter", "meter")` inside of your function call. Look at the summary function for each model and plot the output.

## Instructions part 2

5\) Use `gof_ds()` to assess whether your model fits the data.

6\) Compare the three models using `AIC()`. Which of these is the best model?

7\) Extract your estimated abundance and 95% confidence intervals from each model. Do the confidence intervals contain our "true" estimate of 800 individuals?

------------------------------------------------------------------------

::: panel-tabset
### Detections Figure

```{r, echo = FALSE}
tracklines<-read_sf("data/tracklines.shp")
region<-read_sf("data/region.shp")
ggplot()  + 
  geom_sf(data = region, fill = "lightblue") +  
  geom_sf(data = tracklines, color = "blue") + 
  geom_point(data = bw_dat, aes(x = x, y = y))
```

### Code

```{r, eval = FALSE, echo = TRUE}
tracklines<-read_sf("data/tracklines.shp")
region<-read_sf("data/region.shp")
ggplot()  + 
  geom_sf(data = region, fill = "lightblue") +  
  geom_sf(data = tracklines, color = "blue") + 
  geom_point(data = bw_dat, aes(x = x, y = y))
```

### Distances

```{r, echo = FALSE}
ggplot() + geom_histogram(data = bw_dat, aes(x = distance ), bins = 10)
```

### Code

```{r, eval = FALSE, echo = TRUE}
ggplot() + geom_histogram(data = bw_dat, aes(x = distance ), bins = 10)
```
:::

------------------------------------------------------------------------

## Detection Functions {.scrollable}


### Half-normal

`"hn`"

```{r, echo = FALSE}
d<-seq(1:200)
sigma <- 50
g<-exp((-d^2)/(2*sigma^2))
ggplot() + geom_line(aes(x = d, y = g)) + 
labs(x = "Distance", y = "Detection probability") + 
theme_classic()

```

### Hazard rate

`"hr"`

```{r, echo = FALSE}
d<-seq(1:200)
sigma <- 130
z<-10
g<-(1-exp(-(d/sigma)^(-z)))
ggplot() + geom_line(aes(x = d, y = g)) + 
labs(x = "Distance", y = "Detection probability") + 
theme_classic()
```


### Uniform

`"unif"`

```{r, echo = FALSE}
d<-seq(1:200)
sigma <- 150
z<-0
g<-c(rep(0.5, 150), rep(0, 50))
ggplot() + geom_line(aes(x = d, y = g)) + 
labs(x = "Distance", y = "Detection probability") + 
theme_classic()
```


## Instructions part 2

5\) Use `gof_ds()` to assess whether your model fits the data.

6\) Compare the three models using `AIC()`. Which of these is the best model?

7\) Extract your estimated abundance and 95% confidence intervals from each model. Do the confidence intervals contain our "true" estimate of 800 individuals?

## Discussion
Which model is the best model?

. . .

What might be some challenges with fitting this kind of model? How can we account for these?
